### 2020-06-21 article.tex
Finish working on KSPHPDDM (some TODO left)

% SIAM Article Template
\documentclass[draft,review,onefignum,onetabnum]{siamart190516}

@@

Alongside the paper introducing GCRODR, a MATLAB implementation was provided and
is since then

@@

this paragraph is to explain how the results have been reproduced in PETSc. % Thanks to \pk{KSPHPDDM}, it is quite straightforward. First, the sequence of MATLAB systems had to be converted to PETSc \pk{Mat} and \pk{Vec}. Then, it is just a matter of calling \pk{KSPSolve} in sequence.

@@

    \item in PETSc with \pk{KSPGMRES}, unpreconditioned unrestarted GMRES($\infty$);

@@

All PETSc tests are run on four MPI
processes. They can be launched using the following command line. \\[-4pt]

@@

The various number of iterations needed to reach the convergence criterion as
originally set to $10^{-10}$ are gathered in~\cref{fig:comparison}.
ones from the original GCRODR paper~\cite{parks2006recycling}. Furthermore,
\pk{PCICC} and ichol from MATLAB are not consistent.  Apart from
the tests \ref{pgfplots:r2} and \ref{pgfplots:r3} using $ICC(0)$ and \ref{pgfplots:r4} and \ref{pgfplots:r5}
using the Jacobi method, one may notice that \ref{pgfplots:r0} and
\ref{pgfplots:r1} (resp. \ref{pgfplots:r6} and \ref{pgfplots:r7}) are almost a
perfect match and reproduce partially Figure 4.2 from Parks et
al.~\cite{parks2006recycling}.

@@

        between \ref{pgfplots:r2} and \ref{pgfplots:r3}. All other \pk{KPSHPDDM}
        numbers of iterations match with the respective MATLAB or PETSc
        reference implementation.\label{fig:comparison}}

@@

SLEPc~\cite{hernandez2005ssf}, the Scalable Library for Eigenvalue Problem
computations, is built on top of PETSc. Thus, it is possible to reuse and
compose \pk{KSP} or preconditioners for solving eigenproblems efficiently, e.g.,
through spectral transformations. Recently, it has been observed that in the
field of computational fluid dynamics, the role of recycling can be quite
significant when performing linear stability analysis~\cite{moulin2018al}.
The following generalized eigenvalue problem is considered next:

@@

evaluated using a given base flow $q_b = \begin{bmatrix}u_b & p_b\end{bmatrix}^T$ and $M$ the discretization of the mass matrix
on the space of velocities. There are many ways to compute the eigenvalues
of~\cref{eq:lsa} near a complex-valued shift $\sigma$, but emphasis will be put
in this paragraph on Krylov--Schur methods~\cite{Stewart2002}. These rely on the successive
solution of linear systems such as:
\begin{equation}\label{eq:ks}
    \left(J(q_b) - \sigma \begin{bmatrix}M&0\\0&0\end{bmatrix}\right)x_i = b_i, i = 1,2,\ldots
In SLEPc, these systems are parameterized using the \pk{-st\_} prefix, e.g.,
\pk{-st\_pc\_type~lu}. If they are solved iteratively, it is possible to
use a \pk{KSP} of type \pk{KSPHPDDM} to recycle information between
successive linear systems yielded by the convergence of the Krylov--Schur
algorithm. Using the FreeFEM~\cite{hecht2012new} mini-app from \url{https://github.com/prj-/moulin2019al}
which precondition~\cref{eq:ks} using a modified augmented Lagrangian
approach~\cite{Benzi2011a}, the effect of such recycling is shown next for
finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
flow past a cylinder at Reynolds 100, interested readers are again referred to
the previously cited paper~\cite{moulin2018al} for more details and for larger
runs. Results can be reproduced using the following four commands, the first two
being merely for generating the base flow $q_b$ with \pk{SNESSolve} using a
continuation method on the Reynolds number. \\[-4pt]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 50 -v 0
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 100 -v 0
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
  -st_ksp_type fgmres -st_ksp_gmres_restart 200 -st_ksp_converged_reason
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
  -st_ksp_type hpddm -st_ksp_gmres_restart 200 -st_ksp_converged_reason              &fvtextcolor[myred][\]
  -st_ksp_hpddm_variant flexible -st_ksp_hpddm_recycle 10 -st_ksp_hpddm_type gcrodr
The Krylov--Schur algorithm converges in six outer iterations. In each of these
iterations, multiple linear systems having the structure of~\cref{eq:ks} are
solved. In~\cref{fig:comparison-slepc}, numbers of iterates per inner solves
are reported. After the first inner solve, once recycling kicks in, one may
clearly see the benefits of switching from \pk{KSPFGMRES} to \pk{KSPHPDDM}.
Indeed, through the 46 inner solves, these numbers are roughly decreased by 25. In terms of runtime, \SI[round-mode          = places,round-precision=1]{1.5098e+03}{\sec} (resp.~\SI[round-mode          = places,round-precision=1]{1.0524e+03}{\sec}) is spent in \pk{EPSSolve} when using \pk{KSPFGMRES} (resp.~\pk{KSPHPDDM}). These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}.

@@

As detailed in~\cref{sec:interface-ksp}, it is currently not possible with most
PETSc preconditioners to apply them to a block of vectors. Even though there
are efforts to enable such an operation for an increasing number of \pk{PC}, in
this section, the focus is instead put on the matrix--matrix multiplication
needed in block Krylov methods or when restarting recycling Krylov methods.
In particular, the performance of three PETSc matrix types are evaluated:
    \item \pk{MatSeqAIJ}, standard sequential sparse matrices, based on compressed sparse row format;
    \item \pk{MatSeqBAIJ}, block sparse matrices, based on block compressed sparse row format;
    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrices.
Parallel matrices in PETSc heavily rely on sequential matrices. For example, \pk{MatMPIAIJ}, standard parallel sparse matrices, store on each rank two \pk{MatSeqAIJ}, one with local rows and columns, another with local rows and ``nonlocal'' columns.
To keep this study succinct, only nondistributed sparse matrices will be evaluated next, which can help draw conclusions for the intraprocess performance of parallel sparse matrices. The Seq substring is now dropped, since in this case, \pk{MatSeqAIJ} is equivalent to \pk{MatAIJ} and so on. The performance of the multiplication primitives of the three aforementioned types will also be compared to MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread. There is an ongoing effort to better integrate these routines in PETSc respective types \pk{MatAIJMKL} and \pk{MatBAIJMKL}, but at the time of writing, direct calls to the MKL were used.

The Poisson equation on a cube is discretized with $\mathbb{P}_1$ finite elements in double precision with 32-bit integers. This yields a matrix $A$ of dimension one million. Though the numerical values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered, e.g., when discretizing partial differential equations. Then, a random symmetric dense matrix $T$ of dimension $b$ is generated, for $b \in \{ 1,3,6 \}$, and the following matrix is evaluated $\mathbb{A} = A \otimes T$. The matrix $\mathbb{A}$ is then assembled into all three formats described above. For block matrices, a block size of $b$ is used. The performance of the \pk{MatProductNumeric} operation is evaluated for tall-and-skinny dense matrices with a varying number of columns $N \in \{1,2,8,16,32,64\}$. For $N=1$, the \pk{MatMult} operation for matrix--vector multiplication is used instead. All results are gathered in~\cref{fig:perfAX}. They can be reproduced using the following mini-app: \url{http://jolivet.perso.enseeiht.fr/MatProduct.c} and any input \pk{MatAIJ} stored in binary format, here using the default name \pk{binaryoutput}.

@@

   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types. As baseline values with $N=1$, performance of \pk{MatAIJ} matrix--vector multiplication is used~\ref{pgfplots:aij1}. Results are averages of 5 consecutives multiplications.\label{fig:perfAX}}

@@

    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MatAIJ} is counterproductive, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even with the MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}. However, with this format, as soon as the number of columns is greater or equal than 8, the efficiency reaches approximately 200\%, see~\ref{pgfplots:aij1}; PETSc implementation seems slightly better than MKL except for the single-column case;

@@

    \item for greater block sizes, \texttt{?gemv} or \texttt{?gemm} are used instead for block formats in PETSc. For \pk{MatAIJ}~\ref{pgfplots:aij10} in bottom plot $b=10$, the efficiency quickly caps near 400\% for $N\in\{8,16,32,64\}$, but for block formats, this reaches 500\% for large number of columns, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$. With these numbers of columns, PETSc and MKL perform similarly.

@@

The mini-app may also be used to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or using CUDA.
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][export] OMP_NUM_THREADS=20 &&&& &fvtextcolor[myyellow][export] MKL_NUM_THREADS=20
  -bs 1,3,6 -N 1,2,8,16,32,64 -type aijmkl,baijmkl,sbaijmkl,aijcusparse
Results are reported in~\cref{fig:perfAXpar}. They have been obtained on a single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math mode=true]{261} nodes with two 20-core Intel Xeon Gold 6248 clocked at \SI{2.5}{\giga\hertz} and four NVIDIA Tesla V100 SXM2. Neither transfers between host and device nor time spent in \texttt{mkl\_sparse\_optimize} are taken into account. Once again, it is clear that the larger the number of right-hand sides $N$, the better the efficiency, no matter the block size.

@@

   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc using intranode parallelism. As baseline values with $N=1$, performance of \pk{MatAIJMKL} matrix--vector multiplication is used~\ref{pgfplots:aijmkl1}. Results are averages of 5 consecutives multiplications.\label{fig:perfAXpar}}

@@

>>>>

% SIAM Article Template
\documentclass[review,onefignum,onetabnum]{siamart190516}

@@

Alongside the paper introducing GCRODR~\cite{parks2006recycling},
a MATLAB implementation was provided and is since then

@@

this paragraph is to explain how the results have been reproduced in PETSc.

@@

    \item in PETSc with unpreconditioned and unrestarted GMRES($\infty$) \pk{KSPGMRES};

@@

The notation GCRODR($n$, $m$) indicates a restart after $n$ iterations and a recycling subspace of dimension $m$.
In all the cases, convergence is declared when the initial unpreconditioned residual is reduced by 10 orders of magnitude.
ones from the original GCRODR paper~\cite{parks2006recycling}.
All PETSc tests are performed using four MPI
processes and they can be launched using the following command line: \\[-4pt]

@@

The iteration numbers needed to reach convergence are gathered in~\cref{fig:comparison};
the cases \ref{pgfplots:r0} and
\ref{pgfplots:r1} (resp. \ref{pgfplots:r6} and \ref{pgfplots:r7}) reproduce Figure 4.2 in~\cite{parks2006recycling}.
We note here that the results obtained with
\pk{PCICC} from PETSc (\ref{pgfplots:r3}) and ichol from MATLAB (\ref{pgfplots:r2}, which uses SuiteSparse) are not consistent.
%Apart from
%the tests \ref{pgfplots:r2} and \ref{pgfplots:r3} using $ICC(0)$ and \ref{pgfplots:r4} and \ref{pgfplots:r5}
%using the Jacobi method, one may notice that \ref{pgfplots:r0} and
%\ref{pgfplots:r1} (resp. \ref{pgfplots:r6} and \ref{pgfplots:r7}) are almost a
%perfect match and reproduce partially Figure 4.2 from Parks et
%al.~\cite{parks2006recycling}.

@@

        between \ref{pgfplots:r2} and \ref{pgfplots:r3}. In the other cases,
        the numbers of iterations needed by KSPHPDDM match with the respective
        MATLAB or PETSc reference implementation.\label{fig:comparison}}

@@

In this section we apply recycling Krylov methods in the context linear
stability analysis and consider the solution the following generalized eigenvalue problem:

@@

evaluated using a given base flow $q_b = \begin{bmatrix}u_b & p_b\end{bmatrix}^T$ and
$M$ the discretization of the mass matrix on the space of velocities. Interested
readers are referred to~\cite{moulin2018al} for more details and for larger
runs.

The mini-app employed for the numerical results is built on top of FreeFEM~\cite{hecht2012new} and it is
available at \url{https://github.com/prj-/moulin2019al}. As eigenvalue solver, we consider the Scalable Library for Eigenvalue Problem
computations (SLEPc)~\cite{hernandez2005ssf}, which is built on top of PETSc,
and thus reuse its linear solver infrastructure.

There are different methods to compute the eigenvalues
of~\cref{eq:lsa} near a complex-valued shift $\sigma$; in this paragraph, emphasis will be put
on Krylov--Schur methods~\cite{Stewart2002}, which rely on spectral tranformation steps
and on the solution of successive linear systems such as:
\begin{equation}\label{eq:ks}
    \left(J(q_b) - \sigma \begin{bmatrix}M&0\\0&0\end{bmatrix}\right)x_i = b_i,
\end{equation}
which, in SLEPc, are parameterized using the \pk{-st\_} prefix.
In this work, the above linear system is preconditioned using a modified augmented Lagrangian
approach~\cite{Benzi2011a,moulin2018al}

The effectiveness of subspace recycling is shown for
finding the 5 eigenpairs closest to the shift $\sigma = 10^{-6} + 0.6\mathfrak{i}$ for a
flow past a cylinder at Reynolds 100. Results can be reproduced using the following four commands;
we note here that the first two commands are merely used to generate the base flow $q_b$ with \pk{SNESSolve} using a
continuation method on the Reynolds number. \\[-4pt]
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 50 -v 0
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Nonlinear-solver.edp -Re 100 -v 0
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
  -st_ksp_type fgmres -st_ksp_gmres_restart 200 -st_ksp_converged_reason
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 4 FreeFem++-mpi Eigensolver.edp -Re 100 -v 0 -st_ksp_rtol 1.0e-4         &fvtextcolor[myred][\]
  -st_ksp_type hpddm -st_ksp_gmres_restart 200 -st_ksp_converged_reason              &fvtextcolor[myred][\]
  -st_ksp_hpddm_variant flexible -st_ksp_hpddm_recycle 10 -st_ksp_hpddm_type gcrodr
\end{Verbatim}

The Krylov--Schur algorithm converges in six outer iterations.
Recycling of Krylov subspaces with \pk{KSPHPDDM} provides algorithmic speedup by lowering the number of iterations per inner solve,
as reported in~\cref{fig:comparison-slepc}, from 46 with \pk{KSPFGMRES} to 25.
A similar speedup can be observed in terms of runtime, with the time spent for the solution of the eigenvalue problem being reduced from \SI[round-mode = places,round-precision=1]{1509}{\sec} to \SI[round-mode = places,round-precision=1]{1052}{\sec}. {\color{red} FIXME LATER: These timings, and all which follow except in~\cref{sec:primitives}, have been obtained on {Ir\`ene}, a system composed of \pgfmathprintnumber[assume math mode=true]{1656} nodes with two 24-core Intel Xeon Platinum 8168 clocked at \SI{2.7}{\giga\hertz}}.

@@

We close this section discussing the readiness of PETSc (as of version 3.13.2)
in delivering efficient implementations of the core matrix--matrix multiplication
needed by block Krylov methods and by recycling Krylov methods when restart is performed.

In particular, we consider three PETSc matrix types:
\begin{itemize}
    \item \pk{MatSeqAIJ}, standard sequential sparse matrices, based on compressed sparse row format;
    \item \pk{MatSeqBAIJ}, block sparse matrices, based on block compressed sparse row format;
    \item \pk{MatSeqSBAIJ}, symmetric block sparse matrices stored in upper triangular form.
\end{itemize}
Their distributed memory counterparts rely on the sequential implementations, with each process
storing two such matrices, one with local rows and columns, and another with local rows and ``nonlocal'' columns.
To keep this study succinct, we only evaluate the performances of the sequential implementations, which can help
in drawing conclusions for the intraprocess performance of the parallel formats.
For this reason, from now on we drop the \pk{Seq} substring.
The performances of the multiplication primitives of the three aforementioned types will also be compared to
those obtained with the MKL inspector--executor sparse BLAS routines~\cite{mkl-web-page}, using a single OpenMP thread.
We note that there is an ongoing effort to better integrate these routines in PETSc, but at the time of writing, direct calls to the MKL were used.

To perform the benchmarking, we first discretize the Poisson equation on a cube with $\mathbb{P}_1$ finite elements
and obtain a matrix $A$ of dimension one million. Though the numerical
values of $A$ are of no interest here, the sparsity pattern is rather common and frequently encountered when
discretizing partial differential equations. We then generate a random symmetric dense matrix $T$ of dimension $b$,
for $b \in \{ 1,3,6 \}$, and we construct the matrix $\mathbb{A} = A \otimes T$ into all three formats described above, using
a block size of $b$ for the block storage formats.
The performance of the \pk{MatProductNumeric} operation is evaluated for tall-and-skinny dense matrices with a varying number
of columns $N \in \{2,8,16,32,64\}$. For the single column case $N=1$, the \pk{MatMult} operation is used.
All results have been obtained using double precision arithmetic and 32-bit integers; we report the scaled efficiency (measured in GFLOP per second)
with respect to the baseline \pk{MatAIJ} implementation with $N=1$ in~\cref{fig:perfAX}, where performances have been averaged over five consecutive
product operations.
They can be reproduced by using the mini-app \url{http://jolivet.perso.enseeiht.fr/MatProduct.c} and any input \pk{MatAIJ} stored in binary format,
here using the default name \pk{binaryoutput} {\color{red} Do we want to provide the binary of the matrix used?}.

@@

   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc for three sparse matrix types.\label{fig:perfAX}}

@@

    \item with a block size of 1, top plot $b=1$ in~\cref{fig:perfAX}, using any type but \pk{MatAIJ}~\ref{pgfplots:aij1} is counterproductive when $N>1$, see~\ref{pgfplots:baij1} and~\ref{pgfplots:sbaij1}, even when using MKL, see~\ref{pgfplots:mklbaijseq1} and~\ref{pgfplots:mklsbaijseq1}. With 8 or more columns, the performance of \pk{MatAIJ} plateaus and the efficiency reaches approximately 200\%, while the perfomances of the MKL primitives stagnate at around 175\% see~\ref{pgfplots:aij1};
for the single column case, \pk{MatSBAIJ} and \pk{MatAIJMKL} deliver slightly better performances;

@@

    \item for block sizes larger than 5, block entry multiplication with PETSc block formats is performed using \texttt{?gemv} or \texttt{?gemm} operations. For \pk{MatAIJ}~\ref{pgfplots:aij10} in the bottom plot corresponding to $b=6$, the efficiency quickly caps near 400\% for $N\in\{8,16,32,64\}$, while reaching 500\% for large number of columns for the block formats, see~\ref{pgfplots:baij10} and~\ref{pgfplots:sbaij10} for $N=64$. With these numbers of columns, PETSc and MKL perform similarly.

@@

We then reuse the mini-app to benchmark intranode performance of matrix--matrix multiplications, either with multiple OpenMP threads, or by offloading the operation to a GPU and use the cuSPARSE implementation~\cite{cusparse-web-page}.
\begin{Verbatim}[fontsize=\footnotesize,frame=single,framerule=0.1mm,commandchars=&\[\]]
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][export] OMP_NUM_THREADS=20 &&&& &fvtextcolor[myyellow][export] MKL_NUM_THREADS=20
&fvtextcolor[mygreen][$] &fvtextcolor[myyellow][mpirun] -n 1 ./MatProduct -f binaryoutput -log_view                                 &fvtextcolor[myred][\]
  -bs 1,3,6 -N 1,2,8,16,32,64 -type aijmkl,baijmkl,sbaijmkl,aijcusparse
\end{Verbatim}
Results reported in~\cref{fig:perfAXpar} have been scaled with respect to the baseline implementation \pk{MatAIJMKL} with $N=1$. They have been obtained on a single node of {Jean Zay}, a system composed of \pgfmathprintnumber[assume math mode=true]{261} nodes with two 20-core Intel Xeon Gold 6248 clocked at \SI{2.5}{\giga\hertz} and four NVIDIA Tesla V100 SXM2. Neither transfers between host and device nor time spent in \texttt{mkl\_sparse\_optimize} have been taken into account. {\color{red} FEW MORE SENTENCES, percentage of peak, wait for results with different number of threads}

@@

   \caption{Performance of the matrix--matrix multiplication as implemented in PETSc using intranode parallelism. \label{fig:perfAXpar}}

@@

\bibitem{cusparse-web-page}
{\sc NVIDIA}, {\em {cuSPARSE} web page}.
\newblock
  \url{https://docs.nvidia.com/cuda/cusparse/index.html},
  2020.
